Microlensing as a possible probe of event-horizon structure in quasars

### 2016-07-19 paper1/ms.tex
revised sec 5 in place

The mock lightcurve used a crescent source (as explained in
\S\ref{subsec:crescent}) with parameters
\begin{equation}
   \left(R_p, \frac{R_n}{R_p}, \frac{a}{R_p}, \frac{b}{R_p}\right) =
   (1.31, 0.4, 0, 0.3)
\label{eqn:cp}
\end{equation}
corresponding to $r_{1/2}=1$.  Unit length has 30 pixels on the
magnification map.  The source was carried for 250 pixels along part
of path~I, giving 251 brightness values.  Gaussian noise at the level
of 4\% of the current brightness was then added to give the mock
lightcurve.

\subsection{Model-fitting old results}

Note that the mock lightcurve has three nuisance parameters, namely
the brightness normalisation and the start- and endpoints on path~I.
The two latter can also be thought of as the location of the caustic
with respect to the starting point and the speed of the source.

\begin{figure}
\centering
  \includegraphics[width=0.9\hsize]{figures/cc_forward.eps}
\caption{\label{fig:cc_forward} Mock lightcurve fitted with the
  correct source type and caustic.  The upper panel shows the mock
  data points, and running through them is a seven-parameter best fit:
  these being the crescent parameters $(R_p,R_n,a,b)$ and additionally
  the start and end times and the overall normalisation. The curve has
  $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel shows
  the posterior distributions for $r_{1/2}$ (solid histogram) and
  $R_n/R_p$ (dashed histogram).}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=0.9\hsize]{figures/cc_backward.eps}
\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
  except that the fit uses path~II whereas the mock data still come
  from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
  tolerable fit.}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=0.9\hsize]{figures/gc_forward.eps}
\caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but
  showing the best fit to a Gaussian source.  The histogram shows the
  posterior distribution of $r_{1/2}$ (there is no $R_n$).  The curve
  has $\chi^2=344, P=3\times10^{-6}$ indicating rejection of the
  Gaussian source.}
\end{figure}

This mimics the unavoidable systematic error of not knowing the
caustic exactly.

>>>>

The mock lightcurve used a crescent source (as explained in
\S\ref{subsec:crescent}) with parameters
\begin{equation}
   \left(R_p, \frac{R_n}{R_p}, \frac{a}{R_p}, \frac{b}{R_p}\right) =
   (1.31, 0.4, 0, 0.3)
\label{eqn:cp}
\end{equation}
corresponding to $r_{1/2}=1$.  Unit length has 30 pixels on the
magnification map.  The source was carried for 250 pixels along part
of path~I, giving 251 brightness values.  Gaussian noise at the level
of 4\% of the current brightness was then added to give the mock
lightcurve.  If we assume the brightness measurements are nightly, and
the transverse speed is 600~km/s, $r_{1/2}$ comes to 24~light-minutes,
which is plausible for a quasar.

\subsection{Model-fitting results}

We now proceed to fit the mock lightcurve using MCMC, to recover the
source parameters.

\begin{figure}
\centering
  \includegraphics[width=0.9\hsize]{figures/cc_forward.eps}
\caption{\label{fig:cc_forward} Mock lightcurve fitted with the
  correct source type and caustic.  The upper panel shows the mock
  data points, and running through them is a seven-parameter best fit:
  these being the crescent parameters $(R_p,R_n,a,b)$ and additionally
  the start and end times and the overall normalisation.  (Ideal
  lightcurves for the same crescent parameters appear above in
  Figures~\ref{fig:lightcurve_gauss}, \ref{fig:lightcurve_disk},
  \ref{fig:lightcurve_crescent} and \ref{fig:char_points}.)  The curve
  has $\chi^2=243, P=0.32$ indicating a good fit.  The lower panel
  shows the posterior distributions for $r_{1/2}$ (solid histogram)
  and $R_n/R_p$ (dashed histogram).}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=0.9\hsize]{figures/cc_backward.eps}
\caption{\label{fig:cc_backward} Like Figure~\ref{fig:cc_backward},
  except that the fit uses path~II whereas the mock data still come
  from path~I. The curve has $\chi^2=284, P=0.01$ indicating a
  marginally acceptable fit.}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=0.9\hsize]{figures/gc_forward.eps}
\caption{\label{fig:gc_forward} Like Figure~\ref{fig:cc_backward}, but
  showing the best fit to a Gaussian source.  The histogram shows the
  posterior distribution of $r_{1/2}$ (there is no $R_n$).  It appears
  to show $r_{1/2}$ as accurately recovered, but is not to be trusted,
  because the fit has $\chi^2=344, P=3\times10^{-6}$ indicating
  rejection of the Gaussian source.  Looking closely, the curve is a
  little too shallow near $t=3$, and it is unable to fit the second
  sub-peak near $t=5$.}
\end{figure}

Note that fitting involves three nuisance parameters, namely the
brightness normalisation and the start- and endpoints on the path.
The latter two are equivalent to considering the location of the
caustic with respect to the starting point and the speed of the
source.  The fitting method automatically marginalises over all the
nuisance parameters.

Figure \ref{fig:cc_forward} shows a fit to the mock lightcurve and
recovery of $r_{1/2}$ and $R_n/R_p$.

Figure \ref{fig:cc_backward} shows the result when a different caustic
is used to fit, than was used to generate the mock lightcurve.  This
mimics the unavoidable systematic error of not knowing the caustic
exactly.  The fit is slightly degraded, but still acceptable.

Figure \ref{fig:gc_forward} shows the result of attempting to a
Gaussian source.  There is a small but statistically very significant
misfit because a Gaussian curve cannot reproduce the part of the
lightcurve with the slight flattening and then a second sub-peak
(cf.~Figure~\ref{fig:char_points}) as the dark part of the crescent
crosses the caustic.
